---
title: How confident are we in our classification algorithm?
output: pdf_document
geometry: margin=2in
---

Suppose we have obtained what we believe is a biased coin. The probability of
obtaining heads for this coin is $\theta$, but we don't know the value of
$\theta$. For some threshold value, $\theta_\text{c}$, how confident will we be,
after observing $N$ throws of the coin, that $\theta > \theta_\text{c}$?

We describe our belief that the likelihood of heads, $p_\text{H}$, is some
particular value, $\pi$, as a probability density function; for ease of
calculation, let's assume that it is a beta distribution, 
\begin{equation}
P(p_\text{H} = \pi) = \beta_{\alpha, \beta}(\pi).  
\end{equation}

A plausible belief, prior to throwing the coin, might be to take $\alpha = \beta
= 1$; _i.e._, a flat prior over $[0, 1]$. After observing $N$ throws, our
posterior belief will be described approximately by a beta distribution with
parameters $\alpha = N\theta$ and $\beta = N(1-\theta)$.

For example, suppose $\theta = 1$. The plot shows the cumulative distribution
function for $P(\pi)$ after $1\,000$, $2\,000$, $5\,000$, and $10\,000$
observations. The horizontal, dashed lines show the 80%, 95%, and 99% confidence
thresholds that $p_\text{H}$ exceeds some critical value.

```{r echo=FALSE}
plot(function (theta) pbeta(theta, 10001, 1), xlim = c(0.995, 1.0), ylim = c(0, 0.25), xlab = expression(theta), ylab = expression(P(p[H] < theta)))
curve(pbeta(x, 5001, 1), add = TRUE, col = "red")
curve(pbeta(x, 2001, 1), add = TRUE, col = "darkgreen") 
curve(pbeta(x, 1001, 1), add = TRUE, col = "blue")
confs = c(0.01, 0.05, 0.20)
abline(h = confs, lty = "dashed", col = "darkgrey")
abline(v = qbeta(confs, 1001, 1), lty = "dotted", col = "blue")
abline(v = qbeta(confs, 2001, 1), lty = "dotted", col = "darkgreen")
abline(v = qbeta(confs, 5001, 1), lty = "dotted", col = "red")
abline(v = qbeta(confs, 10001, 1), lty = "dotted", col = "darkgrey")
```

So for example, if one wished to set $\theta_\text{c} = 0.999$, then one would
be 99% confident of this after approximately $5\,000$ observsations and 80%
confident after about $2\,000$ observations. Conversely, if one wished
$\theta_\text{c} = 0.9975$, then one would need just over $1\,000$ observations
to be 95% confident in this fact; and about $2\,000$ observations to be 99%
confident.

Now consider a sequence of observations that we wish to classify as ‘true’ or
‘false.’ We write the ground truth as T or F; and the classification as t or f. 






# Appendix: Further $\beta$-functions

```{r echo=FALSE}
plot(function (theta) pbeta(theta, 9991, 11), xlim = c(0.995, 1.0), ylim = c(0, 1), xlab = expression(theta), ylab = expression(P(p[H] < theta)), main = "Temp: sample plot for theta = 0.999")
curve(pbeta(x, 4996, 6), add = TRUE, col = "red")
curve(pbeta(x, 1999, 3), add = TRUE, col = "darkgreen") 
curve(pbeta(x, 1000, 2), add = TRUE, col = "blue")
confs = c(0.01, 0.05, 0.20)
abline(h = confs, lty = "dashed", col = "darkgrey")
abline(v = qbeta(confs, 1000, 2), lty = "dotted", col = "blue")
abline(v = qbeta(confs, 1999, 3), lty = "dotted", col = "darkgreen")
abline(v = qbeta(confs, 4996, 6), lty = "dotted", col = "red")
abline(v = qbeta(confs, 9991, 11), lty = "dotted", col = "darkgrey")
```





